### 2019-07-10 paper/paper.tex
Add numerical gradient checks

\begin{abstract}
With the rising popularity of deep learning techniques, being able to differentiate through arbitrary functions is becoming more and more important. These arbitrary functions contain a lot of implicit knowledge stored in them which when used in combination with a neural network can significantly speed up its training. In this paper we present a fully differentiable renderer, which is designed to be integrated in any Differentiable Programming (DP) pipeline. We interface our renderer with the deep learning library Flux for use in combination with neural networks. We demonstrate the use of this differentiable renderer in complex rendering tasks and show its use in solving inverse graphics problems.

% \todo{Character count should not exceed 600}

@@

% \let\thefootnote\relax\footnote{* Equal Contribution}

@@

In Computer Graphics, a popular technique for rendering photo realistic images is raytracing. Since ray tracing leverages the properties of the materials of the objects in the scene, a natural extension to the rendering problem would be to extract the exact properties of the materials, lighting, etc. given an image of the scene. However, since ray tracing is non-differentiable at some points and calculating the analytic gradients is a very tedious task, it has been difficult to present a general inverse rendering method. As such there is only one framework in our knowledge, redner\cite{Li:2018:DMC}, which has been able to do so by using analytic gradients.

@@

\begin{figure*}[!htb]
    \centering
    \includegraphics[width=\textwidth]{images/performance-speed.png}
    \caption{Performance Gains on using BVH}
    \label{fig:bvh_perf}
\end{figure*}

\begin{figure*}[!htb]
    \centering
    \includegraphics[width=\textwidth]{images/performance-memory.png}
    \caption{Memory Allocation comparison between rendering with and without BVH}
    \label{fig:bvh_mem}
\end{figure*}

\begin{figure*}[!htb]
    \centering
    \begin{subfigure}[b]{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{images/render/teapot_front.jpg}
        \caption*{}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{images/render/teapot_side.jpg}
        \caption*{}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{images/render/teapot_top.jpg}
        \caption*{}
    \end{subfigure}
    \caption{Utah Teapot Render from three different views. The camera definition shown in Listing~\protect[\ref{lst:example_render}] can be easily modified to generate all these views.}
    \label{fig:teapot_render}
\end{figure*}

\section{Background}
\label{sec:rendering}

Rendering is a technique of generating photorealistic or non photorealistic 2D projections from 3D objects. As such there are several algorithms for rendering complex scenes. One of the most popular techniques for photo realistic rendering is ray tracing. However, for real time rendering algorithms like rasterization is used.

Ray Tracing is a technique in computer graphics for rendering 3D graphics with complex light interactions. In this technique, rays are traced backwards from the eye/camera to the light source(s). In the path, the ray of light can undergo reflection and refraction due to interactions with the objects in its path. This method, however, is very computationally expensive and hence difficult to do in real time.

Rendering being a compute intensive operation is generally done in static languages like C++. This makes it very time expensive to develop the software. Also, these languages lack support of state of the art automatic differentiation tools like Zygote \cite{DBLP:journals/corr/abs-1810-07951}, Jax \cite{jax}, etc which are generally implemented for languages like Julia and Python. As such it is difficult to develop differentiable renderers in those languages and then interface with popular deep learning software.

@@

There are several photo-realistic renderers available which contain a vast amount of implicit knowledge. Differentiation allows such renderers to make use of gradients to learn the inverse mapping from an image to its parameter space. This knowledge can then be used in combination with any machine learning / deep learning model to train them in a fast and efficient manner.

@@

@@

Our approach to differentiable rendering is based on first creating a general purpose renderer and then make use of efficient AD tools for the differentiability part. Hence, at its core RayTracer is a fully featured renderer. It contains functionalities for both raytracing and rasterization.

RayTracer gives users a lot of control to the user over the scene they want to render. The user controls the lighting in the scene, the shape and materials of the objects and the camera configuration. It might be worthwhile to note that we don't make any sort of performance compromise in the forward pass (rendering) to allow gradient computation.

@@

We show the benefits of using Acceleration Structures in Figures~[\ref{fig:bvh_perf}] and [\ref{fig:bvh_mem}]. As can be seen from the plots we not only get a good performance gain (Figure~[\ref{fig:bvh_perf}]) in terms of speed, we also end up allocating much less memory (Figure~[\ref{fig:bvh_mem}] on using Bounding Volume Hierarchy.

\subsection{Rendering the Utah Teapot}

In this section we shall render a very popular model in computer graphics: the Utah Teapot. We simply load the model from a wavefront object (obj) file. Support for other types of files is trivial through the MeshIO.jl package with defines file reader for common mesh object files. Once the scene is loaded the next step is to configure the other elements in the scene - the camera and the light(s). Script~[\ref{lst:example_render}] illustrates the code for rendering the teapot model.

\begin{lstlisting}[caption = {Rendering the Utah Teapot Model},
                   label = {lst:example_render},
                   captionpos = b,
                   language = Julia]
# Screen Size
screen_size = (w = 512, h = 512)

@@

# Render the image
color = raytrace(
    origin,
    direction,
    scene,
    light,
    origin,
    2
)
\end{lstlisting}

@@

@@

As we have mentioned previously, one of the primary motivations of RayTracer.jl is for solving the problem of inverse graphics. Being able to compute gradients wrt any scene parameter, means that we can optimize that parameter. The optimization algorithm is pretty straight forward and is general enough to work for any arbitrary parameter. We describe the algorithm in [\ref{alg:inv_render}].

@@

% We shall provide two examples as a validation for our method. In both these experiments we have used Adam\cite{kingma2014adam} for optimization, but one can use any other method like L-BFGS, BFGS.

@@

\begin{figure*}[!htb]
    \centering
    \begin{subfigure}[b]{0.23\textwidth}
        \centering
        \includegraphics[width=\textwidth]{images/camera/initial_guess_image.png}
        \caption{Image with uncalibrated camera}
        \label{fig:cam_guess}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.23\textwidth}
        \centering
        \includegraphics[width=\textwidth]{images/camera/target_image.png}
        \caption{Target Image to be reconstructed}
        \label{fig:cam_target}
    \end{subfigure}
    \centering
    \begin{subfigure}[b]{0.5\textwidth}
        \centering
        \includegraphics[width=\textwidth, height=100px]{images/camera/iter_120.png}
        \caption{Image obtained after optimization}
    \end{subfigure}
    \caption{Calibration of Camera Parameters to reconstruct Image~[\ref{fig:cam_target}] from Image~[\ref{fig:cam_guess}]}
    \label{fig:cam_invrender}
\end{figure*}

\begin{figure*}[!htb]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\linewidth]{images/light/loss_plot_light.png}
        \caption{Loss Values over the Light Configuration Optimization Process}
        \label{fig:loss_plot_light}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.45\textwidth}
        \begin{subfigure}[b]{0.4\textwidth}
            \centering
            \includegraphics[width=\textwidth]{images/light/original.png}
            \caption{Image to be reconstructed}
            \label{fig:target_light}
        \end{subfigure}
        \hfill
        \begin{subfigure}[b]{0.4\textwidth}
            \centering
            \includegraphics[width=\textwidth]{images/light/initial.png}
            \caption{Initial Guess of Lighting Parameters}
            \label{fig:guess_light}
        \end{subfigure}
        \centering
        \hfill
        \begin{subfigure}[b]{0.4\textwidth}
            \centering
            \includegraphics[width=\textwidth]{images/light/161.png}
            \caption{Image produced after 10 iterations}
        \end{subfigure}
        \hfill
        \begin{subfigure}[b]{0.4\textwidth}
            \centering
            \includegraphics[width=\textwidth]{images/light/501.png}
            \caption{Image with converged parameters}
        \end{subfigure}
    \end{subfigure}
    \caption{Optimization of the lighting conditions to reconstruct Image~[\ref{fig:target_light}] from Image~[\ref{fig:guess_light}]}
    \label{fig:light_position}
\end{figure*}

In this section we showcase our differentiable renderer. Using the following two experiments we validate the functionality of the gradients of our renderer. In both the experiments we make use of the Adam optimizer as described in \cite{kingma2014adam}. We interface the raytracer with Flux to use these optimizers. As an alternative, we have tested the functioning of our package with the optimizers present in Optim.

\subsection{\textbf{Calibration of Camera Parameters}}

In this experiment we start with the image of a rectangle under some configuration of the Camera model. Our aim is to reconstruct the image of this rectangle by modifying the focus and the location of the camera. We assume that all the other parameters of the model, like the light configuration, position of objects, etc. are known apriori.

We follow the standard algorithm as defined in [\ref{alg:inv_render}] for optimizing the parameters. We make an initial guess of the parameters and initialize the camera as:

\begin{lstlisting}[language = Julia]
camera_guess =
    Camera(
        Vec3(5.0, -4.0, -20.0),
        Vec3(0.0, 0.0, 0.0),
        Vec3(0.0, 1.0, 0.0),
        90.0,
        3.0,
        screen_size...
    )
\end{lstlisting}

We use the mean squared difference between the rendered image and the target image as the loss and try to minimize that. The Adam optimizer is initialized with a learning rate of 0.05 and the model is declared to have as converged if the loss falls below 10. Figure~[\ref{fig:cam_invrender}] shows the optimization steps.

\subsection{\textbf{Optimizing the Light Source}}

@@

For this experiment we try to optimize the lighting for a scene containing a tree. We are given the target image with proper lighting conditions. We start with an arbitrary lighting and then iteratively improve the lighting using algorithm~[\ref{alg:inv_render}].

We present the images generated during the optimization process in Figure~[\ref{fig:light_position}].

\section{Conclusion}
In conclusion, we have presented how julia can be leveraged to build differentiable systems. Integrating them with the machine learning pipeline can make an end-to-end differentiable pipeline. Using this pipeline we can define differentiable programming algorithm on it to solve the problem. As expected making the pipeline differentiable allows us to exploit the huge amount of implicit knowledge stored in the system.

>>>>

\begin{abstract}
In this paper we present RayTracer.jl, a renderer in Julia that is fully differentiable using source-to-source Automatic Differentiation (AD). This means that RayTracer not only renders 2D images from 3D scene parameters, but it can be used to optimize for model parameters that generate a target image in a Differentiable Programming (DP) pipeline. We interface our renderer with the deep learning library Flux for use in combination with neural networks. We demonstrate the use of this differentiable renderer in rendering tasks and in solving inverse graphics problems.

@@

\begin{figure*}[!htb]
    \centering
    \begin{subfigure}[c]{0.45\textwidth}
        \includegraphics[height=75mm, width=\textwidth]{images/performance-speed.png}
        \caption{Performance Benchmarks}
        \label{fig:bvh_perf}
    \end{subfigure}
    \hfill
    \begin{subfigure}[c]{0.45\textwidth}
        \includegraphics[height=75mm, width=\textwidth]{images/performance-memory.png}
        \caption{Memory Allocation Benchmarks}
        \label{fig:bvh_mem}
    \end{subfigure}
    \caption{Comparison between scenes rendered with and without BVH}
    \label{fig:bvh}
\end{figure*}

\begin{figure*}[!htb]
    \centering
    \begin{subfigure}[b]{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{images/render/teapot_front.jpg}
        \caption*{}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{images/render/teapot_side.jpg}
        \caption*{}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{images/render/teapot_top.jpg}
        \caption*{}
    \end{subfigure}
    \caption{Utah Teapot Render from three different views. The camera definition shown in Listing~\protect\ref{lst:example_render} can be easily modified to generate all these views.}
    \label{fig:teapot_render}
\end{figure*}

@@

Rendering is a technique of generating photorealistic or non photorealistic 2D projections from 3D objects. As such there are several algorithms for rendering complex scenes. One of the most popular techniques for photo realistic rendering is ray tracing. However, for real time rendering algorithms like rasterization are used.

Ray Tracing is a technique in computer graphics for rendering 3D graphics with complex light interactions. In this technique, rays are traced backwards from the eye/camera to the light source(s). The ray of can undergo reflection and refraction due to interactions with the objects in its path. This technique, however, is very computationally expensive and hence difficult to do in real time. But since ray tracing leverages the properties of the materials of the objects in the scene, a natural extension to the rendering problem would be to extract the exact properties of the materials, lighting, etc. given an image of the scene. Unfortunately ray tracing is non-differentiable at some points and calculating the analytic gradients is a very tedious task. This has made it a very difficult task to present a general inverse rendering method. As such there is only one framework in our knowledge, redner\cite{Li:2018:DMC}, which has been able to do so by using analytic gradients.

Rendering being a computationally expensive operation is generally done in static languages like C++. This makes it very time expensive to develop the software. Also, most languages lack the support of state of the art automatic differentiation tools like Zygote \cite{DBLP:journals/corr/abs-1810-07951}, Jax \cite{jax}, etc which are generally implemented for high level languages like Julia and Python. As such it is difficult to develop differentiable renderers in those languages and then interface with popular deep learning software.

@@

@@

There are several photo-realistic renderers available which contain a vast amount of implicit knowledge. Differentiation allows such renderers to make use of gradients to learn the inverse mapping from an image to its parameter space. This knowledge can then be used in combination with any machine learning / deep learning models to train them in a fast and efficient manner. Experiments in \cite{retinaface} demonstrate the effectiveness of incorporate differentiable renderers in deep learning pipelines by achieving State of the Art results in WIDERFace Hard dataset.

@@

\todo[inline]{Add graphs comparing finite differencing to AD}

@@

Our approach to differentiable rendering is based on first creating a general purpose renderer and then make use of efficient AD tools for the differentiability part. Hence, at its core RayTracer is a fully featured renderer. It contains functionalities for both raytracing and rasterization. Additionally, unlike most prior work in differentiable rendering we do not make performance compromises in the forward pass (rendering) to allow gradient computation.

RayTracer gives users a lot of control to the user over the scene they want to render. The user controls the lighting in the scene, the shape and materials of the objects and the camera configuration.

@@

\todo[inline]{Clarifications and Reproducibility}
We show the benefits of using Acceleration Structures in Figures~\ref{fig:bvh_perf} and \ref{fig:bvh_mem}. As can be seen from the plots we not only get a good performance gain (Figure~\ref{fig:bvh_perf}) in terms of speed, we also end up allocating much less memory (Figure~\ref{fig:bvh_mem} on using Bounding Volume Hierarchy.

\subsection{Rendering the Utah Teapot}
\todo[inline]{Clarity and Grammer}

In this section we shall render a very popular model in computer graphics: the Utah Teapot. We simply load the model from a wavefront object (obj) file. Support for other types of files is trivial through the MeshIO.jl package with defines file reader for common mesh object files. Once the scene is loaded the next step is to configure the other elements in the scene - the camera and the light(s). Script~\ref{lst:example_render} illustrates the code for rendering the teapot model.

\noindent
\begin{minipage}{\linewidth}
\begin{lstlisting}[caption = {Rendering the Utah Teapot Model},
                   label = {lst:example_render},
                   captionpos = b,
                   language = Julia]
# Screen Size
screen_size = (w = 512, h = 512)

@@

# Render the image
color = raytrace(
    origin,
    direction,
    scene,
    light,
    origin,
    2
)
\end{lstlisting}
\end{minipage}

@@

\todo[inline]{More clarity in writing}

@@

As we have mentioned previously, one of the primary motivations of RayTracer.jl is for solving the problem of inverse graphics. Being able to compute gradients wrt any scene parameter, means that we can optimize that parameter. The optimization algorithm is pretty straight forward and is general enough to work for any arbitrary parameter. We describe the algorithm in \ref{alg:inv_render}.

@@

@@

\begin{figure*}[!htb]
    \centering
    \begin{subfigure}[c]{0.26\textwidth}
        \centering
        \includegraphics[width=107px, height=80px]{images/camera/initial_guess_image.png}
        \caption{Image with uncalibrated camera}
        \label{fig:cam_guess}
    \end{subfigure}
    \hfill
    \begin{subfigure}[c]{0.26\textwidth}
        \centering          
        \includegraphics[width=107px, height=80px]{images/camera/target_image.png}
        \caption{Image to be reconstructed}
        \label{fig:cam_target}
    \end{subfigure}
    \centering
    \begin{subfigure}[c]{0.47\textwidth}
        \centering
        \includegraphics[width=\textwidth, height=80px]{images/camera/iter_120.png}
        \caption{Image obtained after optimization}
    \end{subfigure}
    \caption{Calibration of Camera Parameters to reconstruct Image~\ref{fig:cam_target} from Image~\ref{fig:cam_guess}}
    \label{fig:cam_invrender}
\end{figure*}

\begin{figure*}[!htb]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\linewidth]{images/light/loss_plot_light.png}
        \caption{Loss Values over the Light Configuration Optimization Process}
        \label{fig:loss_plot_light}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.45\textwidth}
        \begin{subfigure}[b]{0.4\textwidth}
            \centering
            \includegraphics[width=\textwidth]{images/light/original.png}
            \caption{Image to be reconstructed}
            \label{fig:target_light}
        \end{subfigure}
        \hfill
        \begin{subfigure}[b]{0.4\textwidth}
            \centering
            \includegraphics[width=\textwidth]{images/light/initial.png}
            \caption{Initial Guess of Lighting Parameters}
            \label{fig:guess_light}
        \end{subfigure}
        \centering
        \hfill
        \begin{subfigure}[b]{0.4\textwidth}
            \centering
            \includegraphics[width=\textwidth]{images/light/161.png}
            \caption{Image produced after 10 iterations}
        \end{subfigure}
        \hfill
        \begin{subfigure}[b]{0.4\textwidth}
            \centering
            \includegraphics[width=\textwidth]{images/light/501.png}
            \caption{Image with converged parameters}
        \end{subfigure}
    \end{subfigure}
    \caption{Optimization of the lighting conditions to reconstruct Image~\ref{fig:target_light} from Image~\ref{fig:guess_light}}
    \label{fig:light_position}
\end{figure*}

In this section we showcase our differentiable renderer in some toy inverse rendering problems. Using the following two experiments we demonstrate the use of gradients obtained via AD to recover the camera and lighting parameters fr a scene. In both the experiments we make use of the Adam optimizer as described in \cite{kingma2014adam}. We interface the raytracer with Flux to use these optimizers. As an alternative, we have tested the functioning of our package with the optimizers present in Optim.

\subsection{\textbf{Calibration of Camera Parameters}}
\label{sec:calcam}

In this experiment we start with the image of a rectangle (Listing~\ref{lst:rect_light}) under some configuration of the Camera model (Listing~\ref{lst:cam_opt}). Since RayTracer supports only two primitive shapes - Spheres and Triangles, we need to triangulate the rectangle.

\noindent
\begin{minipage}{\linewidth}
\begin{lstlisting}[caption = {Configuration of the Rectangle for Experiment~\ref{sec:calcam}},
                   label = {lst:rect_light},
                   captionpos = b,
                   language = Julia]
scene = [
    Triangle(
        Vec3( 20.0,  10.0, 0.0),
        Vec3( 20.0, -10.0, 0.0),
        Vec3(-20.0,  10.0, 0.0),
        Material(
            color_diffuse = Vec3(0.0, 1.0, 0.0))),
    Triangle(
        Vec3(-20.0, -10.0, 0.0),
        Vec3( 20.0, -10.0, 0.0),
        Vec3(-20.0,  10.0, 0.0),
        Material(
            color_diffuse = Vec3(0.0, 1.0, 0.0)))
]
\end{lstlisting}
\end{minipage}

\noindent
\begin{minipage}{\linewidth}
\begin{lstlisting}[caption = {Camera Parameters to be Reconstructed},
                   label = {lst:cam_opt},
                   captionpos = b,
                   language = Julia]
camera_target =
    Camera(
        Vec3(0.0, 0.0, -30.0),
        Vec3(0.0, 0.0,   0.0),
        Vec3(0.0, 1.0,   0.0),
        90.0,
        1.0,
        screen_size...
    )
\end{lstlisting}
\end{minipage}

\noindent
\begin{minipage}{\linewidth}
\begin{lstlisting}[caption = {Light Configuration},
                   label = {lst:lgt_param},
                   captionpos = b,
                   language = Julia]
light = PointLight(
    Vec3(1.0, 0.0, 0.0),
    100000.0,
    Vec3(0.0, 0.0, -10.0)
)
\end{lstlisting}
\end{minipage}

Our aim is to reconstruct the image of this rectangle (Figure~\ref{fig:cam_target}) by modifying the focus and the location of the camera. We assume that all the other parameters of the model, like the light configuration (Listing~\ref{lst:lgt_param}), position of objects, etc. are known apriori. We use algorithm~\ref{alg:inv_render} for optimizing the parameters. We make an initial guess of the parameters and initialize the camera (Listing~\ref{lst:cam_initial}).

\noindent
\begin{minipage}{\linewidth}
\begin{lstlisting}[caption = {Initial Guess of the Camera Parameters},
                   label = {lst:cam_initial},
                   captionpos = b,
                   language = Julia]
camera_guess =
    Camera(
        Vec3(5.0, -4.0, -20.0),
        Vec3(0.0,  0.0,   0.0),
        Vec3(0.0,  1.0,   0.0),
        90.0,
        3.0,
        screen_size...
    )
\end{lstlisting}
\end{minipage}

As our loss function, we use the mean squared difference between rendered and target images, each with $300\times400$ pixels having fractional RGB values. We minimize loss with the Adam optimizer, with learning rate 0.05, and declare the model to have converged if loss falls below 10 (where the initial loss is in the order of 600). Figure~\ref{fig:cam_invrender} shows the optimization steps.

\subsection{\textbf{Optimizing the Light Source}}
\todo[inline]{More details to reproduce the results}

@@

For this experiment we try to optimize the lighting for a scene containing a tree. We are given the target image with proper lighting conditions. We start with an arbitrary lighting and then iteratively improve the lighting using algorithm~\ref{alg:inv_render}.

We present the images generated during the optimization process in Figure~\ref{fig:light_position}.

\section{Current Limitations}
\todo[inline]{Write this section highlighting the problem for non differentiable parts}

\section{Conclusion}
\todo[inline]{Needs a rewrite using the comments in the review}
In conclusion, we have presented how julia can be leveraged to build differentiable systems. Integrating them with the machine learning pipeline can make an end-to-end differentiable pipeline. Using this pipeline we can define differentiable programming algorithm on it to solve the problem. As expected making the pipeline differentiable allows us to exploit the huge amount of implicit knowledge stored in the system.
